### 2022-03-21 satfol/satfol.tex
jasmin

\blfootnote{In this work I was the main designer behind all the presented techniques.
Marijn Heule did weekly supervision and provided his knowledge of SAT solving to 
guide the design of all techniques. Jasmin Blanchette found the exact conditions
under which superposition remains complete when predicate elimination rule is added
to the calculus. I also implemented and evaluated all the techniques. 
}

@@

We assume the natural extensions of domain, valuation, interpretation and model
(as defined by Fitting \cite{mf-1996-fol}) from unsorted to many-sorted logic.
We also assume the notion of normal models from Sect.~\ref{sec:pre:fol} (with
corresponding notation) and extend it with the notion of \emph{cannonical
models}: A canonical model $\ourmodel$ is a normal model such that for every
element $d$ in one of $\ourmodel$'s domains there exists a ground term $t$ such
that $\ourmodel$ interprets $t$ as $d$. 
We also make use of Herbrand models and Herbrand's theorem \cite[Sect.~5.4]{mf-1996-fol}.
Canonical models generalize Herbrand
models to first-order logic with equality: If a clause set is satisfiable, then
it is satisfiable in a canonical model.

@@

      % First, $C$ must have the form $\arbpredlit{\cst{p}(\tuple{s}{})} \cor C'$,
      % where $\tuple{s}{}$ is a tuple of ground terms. 
      Without loss of generality,
      we assume that the leading literal of $C$ is positive.
      Towards a contradiction, assume $\xi$ is a valuation such that $\mm \notmodels_\xi C$. Then, $\mm \notmodels_\xi
      \pospredlit{\cst{p}(\tuple{s}{n})}$. 
      % Let $\xi'$ be a valuation that
      % assigns every instance $D = \pospredlit{\cst{p}(\tupleempty{s})} \cor D'$ of a clause in
      % $G$, since $\mm \notmodels \pospredlit{\cst{p}(\tuple{s}{})}$, we have $\mm
      % \models D'$.
      Consider an arbitrary clause $D = \pospredlit{\cst{p}}(\tuplen{x}) \cor D'
      \in \withpredpos{G}{\cst{p}}$ and a valuation $\xi'$, which assigns each $x_i$ the
      interpretation of $s_i$ under $\mm$ and $\xi$. As $\mm \notmodels_{\xi'}
      \pospredlit{\cst{p}}(\tuple{x}{n})$ and $\mm \models G$, then $\mm \models_{\xi'} D'$ for every
      such clause $D$.
   %
      However, by part 5 of Definition~\ref{def:definition} and by Lemma
      \ref{lem:unsat-fresh-consts}, $\mm \notmodels_{\xi'} E(\tuplen{x})$, where $E(\tuplen{x})$
      is the environment associated with the definition set $G$.
      Therefore, there must exist a clause $D =
      \negpredlit{\relax{\cst{p}(\tuplen{x})}} \cor D'$ in $\withpredneg{G}{\cst{p}}$ such
      that $\mm \notmodels_\xi D'$.

@@

\begin{sloppypar}
   The completeness proof follows the invariance-based argument found in the
   article by Waldmann et al \cite{waldmann-et-al-202x-article}.
\end{sloppypar}

@@

In first-order logic, different instances of the same clause can
be used along a transitive chain. For example, the clause $C =
\negpredlit{\cst{p}(x)} \cor \pospredlit{\cst{p}(\cst{f}(x))}$ induces $\cst{p}(x) \bigfimpl{M}^i \cst{p}(\cst{f}^i(x))$ for all $i$.
The algorithm discovers such self-implications (line~\ref{statement:self-impl}): For
each clause~$C$ of the form $\neglit{L} \cor \sigma(L)$, where $\sigma$ is
some nonidempotent substitution, the entires $(\sigma^2(L), \{C\}),\allowbreak
\ldots,\allowbreak (\sigma^{K_{\mathrm{imp}} + 1}(L), \{C\})$ are added to
$\impmap[L]$, where $K_{\mathrm{imp}}$ is a parameter.

>>>>

\blfootnote{In this work I was the main designer of all the presented techniques.
Marijn Heule did weekly supervision and provided his knowledge of SAT solving to 
guide the design of all techniques. Jasmin Blanchette found the exact conditions
under which superposition remains complete when the predicate elimination rule is added
to the calculus. I also implemented and evaluated all the techniques. 
}

@@

We assume the natural extensions of domain, valuation, interpretation, and model
(as defined by Fitting \cite{mf-1996-fol}) from unsorted to many-sorted logic.
We also assume the notion of normal models from Sect.~\ref{sec:pre:fol} (with
corresponding notation) and extend it with the notion of \emph{cannonical
models}: A canonical model $\ourmodel$ is a normal model such that for every
element $d$ in one of $\ourmodel$'s domains there exists a ground term $t$ such
that $\ourmodel$ interprets $t$ as $d$. 
We also make use of Herbrand models and Herbrand's theorem \cite[Sect.~5.4]{mf-1996-fol}.
Canonical models generalize Herbrand
models to first-order logic with equality: If a clause set is satisfiable, then
it is satisfiable by a canonical model.

@@

      % First, $C$ must have the form $\arbpredlit{\cst{p}(\tuple{s}{})} \cor C'$,
      % where $\tuple{s}{}$ is a tuple of ground terms. 
      Without loss of generality,
      we assume that the leading literal of $C$ is positive (i.e., $C$ is of the form $\pospredlit{\cst{p}(\tuple{s}{n})} \cor C'$).
      Towards a contradiction, assume $\xi$ is a valuation such that $\mm \notmodels_\xi C$. Then, $\mm \notmodels_\xi
      \pospredlit{\cst{p}(\tuple{s}{n})}$. 
      % Let $\xi'$ be a valuation that
      % assigns every instance $D = \pospredlit{\cst{p}(\tupleempty{s})} \cor D'$ of a clause in
      % $G$, since $\mm \notmodels \pospredlit{\cst{p}(\tuple{s}{})}$, we have $\mm
      % \models D'$.
      Consider an arbitrary clause $D = \pospredlit{\cst{p}}(\tuplen{x}) \cor D'
      \in \withpredpos{G}{\cst{p}}$ and a valuation $\xi'$, which assigns each $x_i$ the
      interpretation of $s_i$ under $\mm$ and $\xi$. As $\mm \notmodels_{\xi'}
      \pospredlit{\cst{p}}(\tuple{x}{n})$ and $\mm \models G$, then $\mm \models_{\xi'} D'$ for every
      such clause $D$.
   %
      However, by part 5 of Definition~\ref{def:definition} and by Lemma
      \ref{lem:unsat-fresh-consts}, $\mm \notmodels_{\xi'} E(\tuplen{x})$, where $E(\tuplen{x})$
      is the environment associated with the definition set $G$.
      Therefore, there must exist a clause $D =
      \negpredlit{\relax{\cst{p}(\tuplen{x})}} \cor D'$ in $\withpredneg{G}{\cst{p}}$ such
      that $\mm \notmodels_\xi D'$.

@@

\begin{sloppypar}
   The completeness proof follows the invariance-based argument found in the
   article by Waldmann et al. \cite{waldmann-et-al-202x-article}.
\end{sloppypar}

@@

In first-order logic, different instances of the same clause can
be used along a transitive chain. For example, the clause $C =
\negpredlit{\cst{p}(x)} \cor \pospredlit{\cst{p}(\cst{f}(x))}$ induces $\cst{p}(x) \bigfimpl{M}^i \cst{p}(\cst{f}^i(x))$ for all $i$.
The algorithm discovers such self-implications (line~\ref{statement:self-impl}): For
each clause~$C$ of the form $\neglit{L} \cor \sigma(L)$, where $\sigma$ is
some nonidempotent substitution, the pairs $(\sigma^2(L), \{C\}),\allowbreak
\ldots,\allowbreak (\sigma^{K_{\mathrm{imp}} + 1}(L), \{C\})$ are added to
$\impmap[L]$, where $K_{\mathrm{imp}}$ is a parameter.
